---
title: "Soil N2O Emissions-Regression Model"
---

This project contains two parts. The firs part estimates carbon stock change factors (i.e., emission factors) for tillage management and a developed Tier 3 regression model for soil N~2~O emissions through an
analysis of measurement data.  I derived Tier 2 emission factors using simple averages and linear regression models.

### The second part of this project fits the Tier 3 regression model from part 1 based on measured N2O emissions. An R script function was created to estimate emissions from soil carbon stock changes.

Part 1: Emission Factor Development R Script

```{r eval=FALSE}
library(nlme)

###### Management Factor
# Read data from csv file
management.data<-read.csv("SoilCManagement.csv", header=TRUE)

# test for correlation in predictor variables
cor(management.data[,c("years", "dep1", "dep2")])

# fit full model with all variables as main effects for input
test.fit<-lme(ch.cstock~ch.till+years+years2+dep1+dep2+moisture+temp, random=~1 | ran.exp/ran.yrexp, 
              data=management.data, method="ML", na.action=na.omit)
summary(test.fit)

# Diagnostic Plots, Residual Plot
resid<-residuals(test.fit)
plot(fitted(test.fit), resid)
abline(0,0)

# QQ normal plot
qqnorm(resid)
qqline(resid)


# Full model
test.fit<-lme(ch.cstock~ch.till+years+years2+dep1+dep2+moisture+temp, random=~1 | ran.exp/ran.yrexp, 
              data=management.data, method="ML", na.action=na.omit)

# Backward stepwise fit method
test.fit<-lme(ch.cstock~ch.till+years+dep1+dep2+moisture+temp, random=~1 | ran.exp/ran.yrexp, 
              data=management.data, method="ML", na.action=na.omit)

test.fit<-lme(ch.cstock~ch.till+years+dep1+dep2+moisture, random=~1 | ran.exp/ran.yrexp, 
              data=management.data, method="ML", na.action=na.omit)

summary(test.fit)

# Test interactions
test.fit<-lme(ch.cstock~ch.till+years+dep1+dep2+moisture+ ch.till*years+ch.till*dep1+
                ch.till*dep2+ch.till*moisture+years*dep1+years*dep2+years*moisture+
                dep1*moisture+dep2*moisture, random=~1 | ran.exp/ran.yrexp, 
              data=management.data, method="ML", na.action=na.omit)

test.fit<-lme(ch.cstock~ch.till+years+dep1+dep2+moisture+ ch.till*years+ch.till*dep1+
                ch.till*dep2+ch.till*moisture+years*dep1+years*dep2+
                dep1*moisture+dep2*moisture, random=~1 | ran.exp/ran.yrexp, 
              data=management.data, method="ML", na.action=na.omit)

test.fit.management<-lme(ch.cstock~ch.till+years+dep1+dep2+moisture+ ch.till*years+ch.till*dep1+
                ch.till*dep2+years*dep1+years*dep2+
                dep1*moisture+dep2*moisture, random=~1 | ran.exp/ran.yrexp, 
              data=management.data, method="REML", na.action=na.omit)
summary(test.fit.management)


# Derive PDF
fixed.management<-fixed.effects(test.fit.management)
management.cov<-test.fit.management$varFix

# Variables
x.rt.wet<-c(1,1,20,15,300,1,20,15,300,300,6000,15,300)
x.rt.dry<-c(1,1,20,15,300,0,20,15,300,300,6000,0,0)

x.nt.wet<-c(1,0,20,15,300,1,0,0,0,300,6000,15,300)
x.nt.dry<-c(1,0,20,15,300,0,0,0,0,300,6000,0,0)

# Estimates
t(x.rt.wet)%*%fixed.management
t(x.rt.dry)%*%fixed.management
t(x.nt.wet)%*%fixed.management
t(x.nt.dry)%*%fixed.management

# Variance
v.rt.wet<-(t(x.rt.wet)%*%management.cov%*%x.rt.wet)
v.rt.dry<-(t(x.rt.dry)%*%management.cov%*%x.rt.dry)
v.nt.wet<-(t(x.nt.wet)%*%management.cov%*%x.nt.wet)
v.nt.dry<-(t(x.nt.dry)%*%management.cov%*%x.nt.dry)

# Standard Deviation
sqrt(v.rt.wet)
sqrt(v.rt.dry)
sqrt(v.nt.wet)
sqrt(v.nt.dry)

```

### Part 2: Soil N2O Emissions Estimation R Script

```{r eval=FALSE}
"SynFert.N2O.emissions.Regression"<-
  function(mineralN.amount=75, mineralN.amount.sd=5, beta= a, cov.beta=b, MAPPET=1, nreps=10000, iseed=230984, return.option=1)
    # Script developed by Natalie Wiley
    # Originally developed: 3/2/2022
    # Last updated:
    # Script estmates N2O emissions (kg CO2 eq. per ha per yr) from mineral N fertilization
    ### Arguments
    # mineralN.amount      The amount of mineral N fertilizer added to soil (Kg N per HA)
    # mineralN.amount.sd   Standard deviation of the N mineral fertilizer amount
    # MAPPET               Mean annual precipitation to potential evapotranspiration ratio
    # beta                 R object with betas from LME model
    # cov.beta             R object with covriance matrix for betas from LME model
    # nreps                Number of monte carlo simulations
    # iseed                 Initial seed for random draws
    # return.option        1) list object with the emission mean and confidence intervals for N2O emissions, and
    #                      2) the full vector of all Monte Carlo simulations
##
##### Begin Script
  {
##### Set seed
    set.seed(iseed)
##### Check validity of input variables
    # values equal to or greater than 0 are valid
    check.mineralN.amount<-mineralN.amount>=0&mineralN.amount<=880
    if(!check.mineralN.amount){stop("MineralN amount is not valid.")
    } else {cat("NOTE: Mineral N amount is valid")}
    

    check.mineralN.amount.sd<-mineralN.amount.sd>=0
    if(!check.mineralN.amount.sd){stop("MineralN sd is not valid.")
    } else {cat("NOTE: Mineral N sd is valid")}
    
    
    check.MAPPET<-MAPPET>=0.7&MAPPET<=3.3
    if(!check.MAPPET){stop("MAP:PET ratio is not valid.")
    } else {cat("NOTE: MAP:PET is valid")}
    
    # End Validity Checks
    ##
##### Estimate direct N2O emissions using linear mixed effect model
    # Deterministic Calculation
    # estimate emissions and backtransform (since we had to transform our data to be homogenous and use a log transformation, we now have to backtransform a log # into regular units)
    direct.emission.deterministic.ln<-beta%*%t(cbind(1, mineralN.amount,MAPPET))
    direct.emission.deterministic<- (exp(direct.emission.deterministic.ln)) * (44/28) * 298
    
    # Probabilistic calculation
    # Simulate nreps of fertilizer amounts
    mineralN.amount.sim <- rnorm(nreps, mean = mineralN.amount, sd = mineralN.amount.sd)
    
    # Simulate nreps of beta parameters based on LME model
    # determine number of parameters
    numpar<-length(beta) #number of parameters in the model
    
    # computer choleski decomposition 
    M<-t(chol(cov.beta))
    #
    # generate random normals
    z<-matrix(rnorm(nreps*numpar), numpar, nreps)
    
    # produce simulated betas
    sim.beta<-M%*%z+beta
  }

```
